Predicting the Quality of CDRH3 Antibody Loop Structural Models

### 2022-10-05 paper.tex
Fixed i.e. and e.g.

ABodyBuilder is a modelling server that provides the user with a
confidence score for each region (e.g.CDR-H2) of the antibody
model. The given score is the probability that a specific region
(e.g. CDR-H2) will be will be modeled within a specific RMSD
threshold\cite{Leem2016}. Thus, it can be used to obtain an expected
RMSD value for a given probability (default 75\%). For the CDR-H3 this
score is calculated as a function of the loop length.  The confidence
scorer is described as robust, but less accurate in the case of CDR
loops due to the lack of data\cite{Leem2016}. ABLooper also provides a
confidence metric for the CDR-H3 loop model, which is estimated by the
diversity of a set of predicted conformations for the same
loop\cite{Abanades2022}. However, it remains unclear whether a high
prediction diversity score points towards loops with multiple
conformations or a low quality model. Furthermore, it remains unclear
how well the generated diversity score reflects model
quality\cite{Abanades2022}.

@@

After creating the feature dataset, it is pre-processed (cleaning,
scaling, encoding, see methods for details). Structures with a
resolution below 4\AA\ were removed given their low quality. Identical
whole antibody structures were removed from the dataset, while
instances of different antibodies that matched in loop sequence were
not removed. Models of some of these structures with the same loop
sequence differ significantly. The few large RMSD ranges may stem from
low resolution, e.g. the highest datapoint contains a structure with a
resolution of 3.00 \AA.  Residue differences near the loop may also
explain the conformational difference. Some of these structures are
complexed while others are not, which may also affect the loop
structure.

@@

The target data (i.e. RMSD values) are then transformed from numerical
values to nominal values so that they can be used for
classification. In order to define these nominal categories, the total
RMSD range must be divided into categories. This is done either by
creating uniform classes i.e. 1-2\AA, 2-3\AA\ (the optimal size of which
must be determined), etc. or by creating balanced classes. When
creating balanced classes, the upper and lower thresholds of a
category are chosen in such a way that each class contains an equal
number of instances. This approach is chosen to counteract the
skewness of the RMSD distribution. However, this was found to
negatively affect the final model's predictive power. Therefore,
uniform classes were used.

They are also transformed into a set of binary values according to a
list of RMSD thresholds. This is done so that binary models can be
trained, which will predict the probability e.g. that the model's RMSD
is above 2\AA, 2.2\AA, 2.4\AA, and so on. The number of binary
classifiers incorporated into the first layer have a great effect on
the final model, the general trend being that the more binary
classifiers are used, the better the nominal prediction.

@@

\subsubsection{Scaling}
Normalization and Standardization are tested as scaling methods. Both
approaches are greatly influenced by outliers, and such datapoints are
ideally removed for optimal scaling. Here we define outliers as
datapoints that lie over 1.5 times the interquartile range (IQR) below
the first quartile or above the third quartile. The IQR is defined as
the range between quartile 1, i.e. the median of the lower half of the
data, and quartile 3, i.e. the median of the upper half of the
data. However, across all features there are a total of 632 outlier
values and removing such a large number of datapoints is not a viable
option. A robust scaler was also used, which uses statistics that are
robust to outliers. The median is set to zero and numerical features
are scaled to the interquartile range.

@@

Feature encoding methods that were very high-dimensional
(one-hot-encoding, BLOSUM62, NLF) were found to be
unfavorable. Dimensionality reduction methods (Principal Component
Analysis (PCA), Independent Component Analysis (ICA), projection-based
methods e.g. t-SNE) were used on BLOSUM62 encoded matrices, which lead
to significant improvement. However, a physiochemical encoding
strategy was most effective.

>>>>

ABodyBuilder is a modelling server that provides the user with a
confidence score for each region (e.g.\ CDR-H2) of the antibody
model. The given score is the probability that a specific region
(e.g.\ CDR-H2) will be will be modeled within a specific RMSD
threshold\cite{Leem2016}. Thus, it can be used to obtain an expected
RMSD value for a given probability (default 75\%). For the CDR-H3 this
score is calculated as a function of the loop length.  The confidence
scorer is described as robust, but less accurate in the case of CDR
loops due to the lack of data\cite{Leem2016}. ABLooper also provides a
confidence metric for the CDR-H3 loop model, which is estimated by the
diversity of a set of predicted conformations for the same
loop\cite{Abanades2022}. However, it remains unclear whether a high
prediction diversity score points towards loops with multiple
conformations or a low quality model. Furthermore, it remains unclear
how well the generated diversity score reflects model
quality\cite{Abanades2022}.

@@

After creating the feature dataset, it is pre-processed (cleaning,
scaling, encoding, see methods for details). Structures with a
resolution below 4\AA\ were removed given their low quality. Identical
whole antibody structures were removed from the dataset, while
instances of different antibodies that matched in loop sequence were
not removed. Models of some of these structures with the same loop
sequence differ significantly. The few large RMSD ranges may stem from
low resolution, e.g.\ the highest datapoint contains a structure with a
resolution of 3.00 \AA.  Residue differences near the loop may also
explain the conformational difference. Some of these structures are
complexed while others are not, which may also affect the loop
structure.

@@

The target data (i.e.\ RMSD values) are then transformed from numerical
values to nominal values so that they can be used for
classification. In order to define these nominal categories, the total
RMSD range must be divided into categories. This is done either by
creating uniform classes i.e.\ 1-2\AA, 2-3\AA\ (the optimal size of which
must be determined), etc. or by creating balanced classes. When
creating balanced classes, the upper and lower thresholds of a
category are chosen in such a way that each class contains an equal
number of instances. This approach is chosen to counteract the
skewness of the RMSD distribution. However, this was found to
negatively affect the final model's predictive power. Therefore,
uniform classes were used.

They are also transformed into a set of binary values according to a
list of RMSD thresholds. This is done so that binary models can be
trained, which will predict the probability e.g.\ that the model's RMSD
is above 2\AA, 2.2\AA, 2.4\AA, and so on. The number of binary
classifiers incorporated into the first layer have a great effect on
the final model, the general trend being that the more binary
classifiers are used, the better the nominal prediction.

@@

\subsubsection{Scaling}
Normalization and Standardization are tested as scaling methods. Both
approaches are greatly influenced by outliers, and such datapoints are
ideally removed for optimal scaling. Here we define outliers as
datapoints that lie over 1.5 times the interquartile range (IQR) below
the first quartile or above the third quartile. The IQR is defined as
the range between quartile 1, i.e.\ the median of the lower half of the
data, and quartile 3, i.e.\ the median of the upper half of the
data. However, across all features there are a total of 632 outlier
values and removing such a large number of datapoints is not a viable
option. A robust scaler was also used, which uses statistics that are
robust to outliers. The median is set to zero and numerical features
are scaled to the interquartile range.

@@

Feature encoding methods that were very high-dimensional
(one-hot-encoding, BLOSUM62, NLF) were found to be
unfavorable. Dimensionality reduction methods (Principal Component
Analysis (PCA), Independent Component Analysis (ICA), projection-based
methods e.g.\ t-SNE) were used on BLOSUM62 encoded matrices, which lead
to significant improvement. However, a physiochemical encoding
strategy was most effective.
